### 2021-08-05 src/paper/abstract.tex
Headline, first batch of changes.

\noindent
In order to slow the spread of the CoViD-19 pandemic, governments around the world have
enacted a wide set of policies limiting the transmission of the disease. Initially, these
focused on non-pharmaceutical interventions; more recently, vaccinations and large-scale
rapid testing have started to play a major role. The objective of this study is to
explain the quantitative effects of these policies on determining the course of the
pandemic, allowing for factors like seasonality or virus strains with different
transmission profiles. To do so, the study develops an agent-based simulation model,
which is estimated using data for the second and the third wave of the CoViD-19 pandemic
in Germany. The paper finds that during a period where vaccination rates rose from 5\% to
40\%, seasonality and rapid testing had the largest effect on reducing infection numbers.
Frequent large-scale rapid testing should remain part of strategies to contain CoViD-19;
it can substitute for many non-pharmaceutical interventions that come at a much larger
cost to individuals, society, and the economy.

>>>>

\noindent
In order to slow the spread of the CoViD-19 pandemic, governments around the world have
enacted a wide set of policies limiting the transmission of the disease. Initially,
these focused on non-pharmaceutical interventions; more recently, vaccinations and
large-scale rapid testing have started to play a major role. The objective of this study
is to explain the quantitative effects of these policies on determining the course of
the pandemic, allowing for factors like seasonality or virus strains with different
transmission profiles. To do so, the study develops an agent-based simulation model,
which explicitly takes into account test demand and behavioral changes following
positive tests. is estimated using data for the second and the third wave of the
CoViD-19 pandemic in Germany. The paper finds that during a period where vaccination
rates rose from 5\% to 40\%, seasonality and rapid testing had the largest effect on
reducing infection numbers. Frequent large-scale rapid testing should remain part of
strategies to contain CoViD-19; it can substitute for many non-pharmaceutical
interventions that come at a much larger cost to individuals, society, and the economy.

### 2021-08-05 src/paper/report.tex
Headline, first batch of changes.

At the core of our agent-based model \citep[][we review more literature in Supplementary
Material~\ref{sec:literature_review}]{Aleta2020,Hinch2020} are physical contacts between
heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
an infectious individual and somebody susceptible to the disease bears the risk of
transmitting the virus. Contacts occur in up to four networks: Within the household, at
work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
depend on age and geographical location (see Supplementary
Material~\ref{subsec:assortativity}).

@@

The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
never use it directly. For legibility reasons, we transform the index so that lower
values represent higher levels of restrictions. A value of zero means all measures
incorporated in the index are turned on. The value one represents the situation in
mid-September, with restrictions on gatherings and public events, masking requirements,
but open schools and workplaces. In the seven weeks between mid September and early
November, cases increased by a factor of ten. Restrictions were somewhat tightened in
mid-October and again in early November. New infections remained constant throughout
November before rising again in December, prompting the most stringent lockdown to this
date. Schools and daycare centers were closed, so were customer-facing businesses except
for grocery and drug stores. From the peak of the second wave just before Christmas
until the trough in mid-February, newly detected cases decreased by almost three
quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
    (Delta) reached Germany in April but still accounted for less than 5\% of cases at the
    end of our simulation period.} In early March, some NPIs were relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the public. There
were many changes in details of regulations afterwards, but they did not change the
overall stringency index.

@@

@@

Two of the most contentious NPIs concern schools and mandates to work from home. In many
countries, schools switched to remote instruction during the first wave, so did Germany.
After the summer break, they were operating at full capacity with increased hygiene
measures, before being closed again from mid-December onward. Some states started opening
them gradually in late February, but operation at normal capacity did not resume until
the beginning of June. Figure~\ref{fig:school_scenarios} shows the effects of different
policies regarding schools starting after Easter, at which point rapid tests had become
widely available. We estimate the realized scenario to have essentially the same effect
as a situation with closed schools. Under fully opened schools with mandatory tests,
total infections would have been 6\% higher; this number rises to 20\% without tests.
These effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of the
large negative effects school closures have on children and parents \citep{Luijten2021,
Melegari2021}---and in particular on those with low socio-economic status---these results
in conjunction with hindsight bias suggest that opening schools combined with a testing
strategy would have been beneficial. In other situations, and particular when rapid test
are not available at scale, trade-offs may well be different.

>>>>

\section{Model description}

At the core of our agent-based model \citep[][we review more literature in Supplementary
Material~\ref{sec:literature_review}]{Aleta2020,Hinch2021} are physical contacts between
heterogeneous agents (Figure~\ref{fig:model_contacts_infections}). Each contact between
an infectious individual and somebody susceptible to the disease bears the risk of
transmitting the virus. Contacts occur in up to four networks: Within the household, at
work, at school, or in other settings (leisure activities, grocery shopping, medical
appointments, etc.). Some contacts recur regularly, others occur at random. Empirical
applications can take the population and household structure from census data and the
network-specific frequencies of contacts from diary data measuring contacts before the
pandemic \citep[e.g.][]{Mossong2008,Hoang2019}. Within each network, meeting frequencies
depend on age and geographical location (see Supplementary
Material~\ref{subsec:assortativity}).

@@

\section{Second and third waves of the CoViD-19 pandemic in Germany}

The model is applied to the second and third waves of the CoViD-19 pandemic in Germany,
covering the period mid-September 2020 to the end of May 2021.
Figure~\ref{fig:pandemic_drivers_model_fit} describes the evolution of the pandemic and
of its drivers. The black line in Figure~\ref{fig:aggregated_fit} shows officially
recorded cases; the black line in Figure~\ref{fig:stringency_infectious_contacts} the
Oxford Response Stringency Index \citep{Hale2020}, which tracks the tightness of
non-pharmaceutical interventions. The index is shown for illustration of the NPIs, we
never use it directly. For legibility reasons, we transform the index so that lower
values represent higher levels of restrictions. A value of zero means all measures
incorporated in the index are turned on. The value one represents the situation in
mid-September, with restrictions on gatherings and public events, masking requirements,
but open schools and workplaces. In the seven weeks between mid September and early
November, cases increased by a factor of ten. Restrictions were somewhat tightened in
mid-October and again in early November. New infections remained constant throughout
November before rising again in December, prompting the most stringent lockdown to this
date. Schools and daycare centers were closed, so were customer-facing businesses except
for grocery and drug stores. From the peak of the second wave just before Christmas
until the trough in mid-February, newly detected cases decreased by almost three
quarters. The third wave in the spring of 2021 is associated with the B.1.1.7 (Alpha)
strain, which became dominant in March (Figure~\ref{fig:share_b117}).\footnote{B.1.617.2
(Delta) was first detected in Germany in April; at the end of our simulation period it
accounted for less than 5\% of cases.} In early March, some NPIs were relaxed; e.g.,
hairdressers and home improvement stores were allowed to open again to the public. There
were many changes in details of regulations afterwards, but they did not change the
overall stringency index.

@@

\section{Results}

@@

\section{Discussion and conclusions}

Having quantified the effects of various mechanisms, we now simulate hypothetical
scenarios comparing changes in NPIs and testing regimes. Two of the most contentious
NPIs concern schools and mandates to work from home. In many countries, schools switched
to remote instruction during the first wave, so did Germany. After the summer break,
they were operating at full capacity with increased hygiene measures, before being
closed again from mid-December onward. Some states started opening them gradually in
late February, but operation at normal capacity did not resume until the beginning of
June. Figure~\ref{fig:school_scenarios} shows the effects of different policies
regarding schools starting after Easter, at which point rapid tests had become widely
available. We estimate the realized scenario to have essentially the same effect as a
situation with closed schools. Under fully opened schools with mandatory tests, total
infections would have been 6\% higher; this number rises to 20\% without tests. These
effect sizes are broadly in line with empirical studies (e.g. \citet{Vlachos2021,
Berger2021}, see Section~\ref{subsec:model_validation} for a comparison). In light of
the large negative effects school closures have on children and parents
\citep{Luijten2021, Melegari2021}---and in particular on those with low socio-economic
status---these results in conjunction with hindsight bias suggest that opening schools
combined with a testing strategy would have been beneficial. In other situations, and
particular when rapid test are not available at scale, trade-offs may well be different.
